Douglas-complex fire smoke and cardiopulmonary morbidity

### 2017-11-14 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
fixed table 1 function to handle 0 events in under 15 subgroup for certain cvd outcomes.

```{r health data import, message=F, warning=F}
# importing case-crossover dataframes 
# SABA case-crossover will be imported first
# i can't find a lot of the files
# asthma
asthma_path <- paste0("./data/health/2017-oregon_casecross_asthma.csv")
asthma_df <- read_csv(asthma_path)

# cvd 
cvd_path <- paste0("./data/health/2017-oregon_casecross_cvd.csv")
cvd_df <- read_csv(cvd_path)
# had a warning message for cvd dataframe import; should check on this

# import all other dataframes

# I think purrr would be a good way to loop through a list of outcome dataframes
# bind dataframes in a list
outcome_df_list <- list(asthma_df, cvd_df)

# name list elements in order I'd like to present in document
# commented out for now.
# outcome_names <- c("SABA", "Respiratory", "Asthma", "COPD", "Pneumonia", 
#   "Acute Bronchitis", "Cardiovascular Disease", "Arrhythmia", 
#   "Cerebrovascular Disease", "Heart Failure", "Ischemic Heart Disease", 
#   "Myocardial Infarction")
outcome_names <- c("Asthma", "CVD")

names(outcome_df_list) <- c( 'Asthma', 'CVD')
```

@@

@@

  # output row of summary numbers
  age_cat_n <- data_list %>% 
    filter(outcome == 1) %>%
    group_by(age_ind) %>% 
    summarise(n = n()) %>%
    mutate(age_cat = ifelse(age_ind == 0, "age_under15",
                      ifelse(age_ind == 1, "age_15to65", "age_over65")),
           # preserve factor order
           age_cat = parse_factor(age_cat, levels = age_cat)) %>% 
    select(-age_ind) %>% 
    spread(age_cat, n)
  # sex n
  sex_n <- data_list %>% 
    filter(outcome == 1) %>% 
    group_by(gender) %>% 
    summarise(n = n()) %>% 
    filter(gender != "U") %>% 
    spread(gender, n)
  # bind all rows together
  outcome_n_row <- bind_cols(total_n, age_cat_n, sex_n) %>% 
    mutate(perc_15 = round(age_under15/total_n,2), 
           perc_15to65 = round(age_15to65/total_n,2),
           perc_65 = round(age_over65/total_n,2),
           perc_F = round(F/total_n,2),
           perc_M = round(M/total_n,2)) %>% 
    select(total_n, age_under15, perc_15, age_15to65, perc_15to65, 
           age_over65, perc_65, F, perc_F, M, perc_M)
return(outcome_n_row)
} # end function

# outcome names
outcome_names <- names(outcome_df_list) %>% 
  data_frame() %>% 
  rename(outcome = ".")

# apply custom function to generate descriptive table
table_vals <- outcome_df_list %>% 
  # map to dataframe
  map_dfr(table1_fun) 
# descriptive table, bind outcome names and values
descriptive_table <- bind_cols(outcome_names, table_vals)

@@

```{r cumulative association}
# use custom function to create a dataframe of cumulative values
# using purrr::map_dfr function
# purrr::map_dfr will work if it's just the cumulative, but won't work for lagged
cumulative_results <- outcome_df_list %>% 
  map_dfr(~distributed_lag(., pm_df=3, temp_df=2, cumulative=T)) %>% 
  bind_cols(., data_frame(outcome_names)) %>% 
  select(outcome_names, type, estimate, lower95, upper95)

@@

# ggplot ----
plot <- ggplot(cumulative_results, aes(x=outcome_names, y = estimate)) +
  geom_point(colour = "darkblue") +
  geom_text(aes(label = round(estimate,2)), hjust = -0.25) +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), colour= "darkblue", width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank()) +
  ylab(expression("Cumulative odds ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
  xlab("Outcomes")

>>>>

*Note 2017-11-14: It may be easiest to make a set of casecross over functions and put them in a package rather than import the timestratified casecross overs each time.*

```{r health data import, message=F, warning=F}
# path for health dataframe
health_path <- "./data/health/"
# importing time-stratified case-crossover dataframe
casecross_files <- list.files(path=health_path, pattern="*casecross*")

# reorder file names in way I'd like it to be analyzed
# also ditching broken arm
casecross_order <- casecross_files[c(13,12,3,6,11,1,7,2,5,8,9,10)]

# extract the name of the dataframe after study type
file_names <- stringr::str_extract(casecross_order, 
  pattern="(?<=casecross_)\\w+")

# read in all the files in to a list
outcome_df_list <- casecross_order %>%  
  purrr::map(~read_csv(paste0(health_path, .)))

# name dataframes
names(outcome_df_list) <- file_names

# name list elements in order I'd like to present in document
outcome_names <- c("SABA", "Respiratory", "Asthma", "COPD", "Pneumonia", 
  "Acute Bronchitis", "Cardiovascular Disease", "Arrhythmia", 
  "Cerebrovascular Disease", "Heart Failure", "Ischemic Heart Disease", 
  "Myocardial Infarction")
```

@@

*Also warnings on mixed data types of NULL and numeric; another reason to revise and create timestratified casecross over functions*.

@@

  # output row of summary numbers
  age_cat_n <- data_list %>% 
    filter(outcome == 1) %>%
    group_by(age_ind) %>% 
    summarise(n = n()) %>%
    # this will fix the 0 categories of certain CVD outcomes in children
    add_row(age_ind = 0, n = 0) %>% 
    arrange(age_ind) %>% 
    group_by(age_ind) %>% 
    summarise(n = sum(n)) %>% 
    mutate(age_cat = ifelse(age_ind == 0, "age_under15",
                    ifelse(age_ind == 1, "age_15to65", "age_over65")),
           # preserve factor order
           age_cat = parse_factor(age_cat, levels = age_cat)) %>% 
    select(-age_ind) %>% 
    spread(age_cat, n)
  # sex n
  sex_n <- data_list %>% 
    filter(outcome == 1) %>% 
    group_by(gender) %>% 
    summarise(n = n()) %>% 
    filter(gender != "U") %>% 
    spread(gender, n)
  # bind all rows together
  outcome_n_row <- bind_cols(total_n, age_cat_n, sex_n) %>% 
    mutate(perc_15 = round(age_under15/total_n,2), 
      perc_15to65 = round(age_15to65/total_n,2),
      perc_65 = round(age_over65/total_n,2),
      perc_F = round(F/total_n,2),
      perc_M = round(M/total_n,2)) %>% 
    select(total_n, age_under15, perc_15, age_15to65, perc_15to65, 
           age_over65, perc_65, F, perc_F, M, perc_M)
return(outcome_n_row)
} # end function

# outcome names
outcome <- outcome_names %>% 
  as_tibble() %>%
  rename(outcome = value)

# apply custom function to generate descriptive table
table_vals <- outcome_df_list %>% 
  # map to dataframe
  map_dfr(table1_fun) 
# descriptive table, bind outcome names and values
descriptive_table <- bind_cols(outcome, table_vals)

@@

```{r cumulative association}
# use custom function to create a dataframe of cumulative values
# using purrr::map_dfr function
# purrr::map_dfr will work if it's just the cumulative, but won't work for lagged
cumulative_results <- outcome_df_list %>% 
  map_dfr(~distributed_lag(., pm_df=3, temp_df=2, cumulative=T)) %>% 
  bind_cols(., data_frame(outcome_names)) %>% 
  select(outcome_names, type, estimate, lower95, upper95) %>% 
  # preserve order of names
  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))

@@

# ggplot ----
plot <- ggplot(cumulative_results, aes(x=outcome_names, y = estimate)) +
  geom_point(colour = "darkblue") +
  geom_text(aes(label = round(estimate,2)), hjust = -0.25) +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                colour= "darkblue", width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle=90, hjust = 1)) +
  ylab(expression(paste0("Cumulative odds ratio for a 10 ug/m"^3, 
    " increase smoke PM"[2.5]))) +
  xlab("Outcomes")
